---
title: "2023 FIRE198 Sustainability Analytics Week 4 In-class Activity"
format: html
editor: visual
---

Step 1: Declare that you will use the **tidyverse** package.

```{r}
library("tidyverse")
```

Step 2: Upload **nayro.csv** into the RStudio environment.

```{r}
df<-read.csv("nayro.csv")
```

Step 3: Create a new dataframe that contains only data in Prince George's County, Maryland.

```{r}
glimpse(df)
```

```{r}
#Option 1
df2<-df %>%
  filter(state=="MARYLAND", countyname=="Prince George's County")

#Option 2
df2<-df %>%
  filter(state=="MARYLAND" & countyname=="Prince George's County")

#Option 3
df2<-df %>%
  filter(state=="MARYLAND") %>%
  filter(countyname=="Prince George's County")
```

Step 4: Create a new dataframe that contains data for **both** Prince George's County and Montgomery County in Maryland.

```{r}

```

Step 5: Create a new dataframe containing data from **both** the county **you and the person next to you** are from. Did any pollutant exceed the national standard in 2022?

```{r}

```

Step 6: How many counties exceed that national standard for **PM-2.5 (2006)** in the year **2022?**

```{r}
df_pm<-df %>%
  filter(pollutant=="PM-2.5 (2006)") %>%
  filter(!is.na(yr2022))
```

Step 7: Create a new column that shows the total number of people living in counties that exceeded the national standard for **PM-2.5 (2006)** in **2022?**

```{r}
df_pm2<-df_pm %>%
  mutate(total_pop=sum(population))
```

Step 8: Create a new column in the dataframe **df_pm2** to indicate whether the column class is serious.

Name the new column s**erious_num**.

Let **serious_num** equal to 1 if the column class is equal to **Serious**.

Let **serious_num** equal 0 if the column class is not equal to **Serious**.

```{r}
df_pm3<-df_pm2 %>%
  mutate(serious=ifelse(class=="Serious",1,0))
```
